Enriching Object Detection with 2D-3D Registration and Continuous Viewpoint Estimation

### 2015-04-07 2349.tex
extended abstract and Michael's edit

\author{Christopher Bongsoo Choy\textsuperscript{\dag}, Michael Stark\textsuperscript{\dag \ddag}, Sam Corbett-Davies\textsuperscript{\dag}, Silvio Savarese\textsuperscript{\dag}\\
    \textsuperscript{\dag}Stanford University, \textsuperscript{\ddag}Max Plank Institutes\\
    {\tt\small \{chrischoy, scorbett, stark, ssilvio\}@stanford.edu}
% For a paper whose authors are all at the same institution,
% omit the following lines up until the closing ``}''.
% Additional authors and addresses can be added with ``\and'',
% just like the second author.
% To save space, use either the email address or home page, not both
}

@@

We propose an efficient method for synthesizing templates from 3D models that
runs on the fly -- that is, it quickly produces detectors for an arbitrary
viewpoint of a 3D model without expensive dataset-dependent training or template
storage. Given a 3D model and an arbitrary continuous detection viewpoint, our
method synthesizes a discriminative template by extracting features from a
rendered view of the object and decorrelating spatial dependences among the
features. Our decorrelation procedure relies on a gradient-based algorithm that
is more numerically stable than standard decomposition-based procedures, and we
efficiently search for candidate detections by computing FFT-based template
convolutions. Due to the speed of our template synthesis procedure, we are able
to perform joint optimization of scale, translation, continuous rotation, and
focal length. We provide an efficient GPU implementation of our algorithm, and
we validate its performance on 3D Object Classes and PASCAL3D+ datasets.

@@

\section{Acknowledgement}
This work has been supported by the Max Planck Center for Visual Computing \& Communication.

>>>>

\author{Christopher Bongsoo Choy\textsuperscript{\dag}, Michael Stark\textsuperscript{\dag \ddag}, Sam Corbett-Davies\textsuperscript{\dag}, Silvio Savarese\textsuperscript{\dag}\\
	\textsuperscript{\dag}Stanford University, \textsuperscript{\ddag}Max Planck Institute for Informatics\\
	{\tt\small \{chrischoy, scorbett, ssilvio\}@stanford.edu, \textsuperscript{\ddag}stark@mpi-inf.mpg.de }
% For a paper whose authors are all at the same institution,
% omit the following lines up until the closing ``}''.
% Additional authors and addresses can be added with ``\and'',
% just like the second author.
% To save space, use either the email address or home page, not both
}

@@

We propose an efficient method for synthesizing templates from 3D models that
runs on the fly -- that is, it quickly produces detectors for an arbitrary
viewpoint of a 3D model without expensive dataset-dependent training or template
storage. Given a 3D model and an arbitrary continuous detection viewpoint, our
method synthesizes a discriminative template by extracting features from a
rendered view of the object and decorrelating spatial dependences among the
features. Our decorrelation procedure relies on a gradient-based algorithm that
is more numerically stable than standard decomposition-based procedures, and we
efficiently search for candidate detections by computing FFT-based template
convolutions. Due to the speed of our template synthesis procedure, we are able
to perform joint optimization of scale, translation, continuous rotation, and
focal length using Metropolis-Hastings algorithm. We provide an efficient GPU
implementation of our algorithm, and we validate its performance on 3D Object
Classes and PASCAL3D+ datasets.

@@

\paragraph{Acknowledgement}
This work has been supported by the Max Planck Center for Visual Computing \& Communication.

### 2015-04-07 2349_ext.tex
extended abstract and Michael's edit

% Use class option [extendedabs] to prepare the 1-page extended abstract.
\documentclass[extendedabs]{bmvc2k}
\usepackage[colorlinks = true,
            linkcolor = blue,
            urlcolor  = blue,
            citecolor = blue,
            anchorcolor = blue]{hyperref}

@@

\title{A Plumbline Constraint for the Rational Function Lens Distortion Model}
\addauthor{
David Claus$^1$, 
Andrew Fitzgibbon$^2$
}{}{1}
\addinstitution{
$^1$Department of Engineering Science, Oxford University. 
$^2$Microsoft Research Cambridge.
}

@@

\noindent
Calibration of lens distortion has traditionally proceeded in one
of three ways.  The first is to use known correspondences between
feature points in one or more images, and the world 3D points.
This is typically done using a checkerboard or a calibration grid
of dots where corners or dot centres can be reliably
located~\cite{Tsai86}. A second class of calibration techniques is
termed auto-calibration as it relies solely on detecting static
points within a scene~\cite{Zhang96,Fitzgibbon01}.  This paper fits
into the third category, where straight lines in the world are used
to determine the distortion parameters.  This {\em plumbline}
technique was first mooted by~\cite{Brown71}, and has been applied
to various distortion models~\cite{Devernay01,Swaminathan00}.

We show in this paper how a plumbline constraint can be implemented
using the rational function model for lens
distortion~\cite{Claus05}, under which straight lines are imaged as
conics, and this permits an elegant factorization of the conics
into the camera calibration and the equations of the straight
lines.  This differs from previous plumbline work in two ways:
first, a factorization-based algorithm can be formulated to
estimate the distortion; second, nonlinear refinement of the
distortion can be easily implemented to minimize a good
approximation of geometric distance in the image plane.  The simplicity of the mapping leads to fast and efficient convergence of the
nonlinear algorithm over a range of starting positions.

\newcommand{\tr}{^{\top}}
\newcommand{\matx}[1]{\mbox{\tt #1}}
\newcommand{\vect}[1]{{\bf #1}}
For a perspective camera, the mapping from image pixels $(i,j)$ to 3D rays
$\vect d(i,j)$ can be expressed as:
\begin{equation}
\vect d(i,j) = \begin{pmatrix}
B_{11} i   + B_{12}  j + B_{13} 1\\
B_{21} i   + B_{22}  j + B_{23} 1\\
B_{31} i   + B_{32}  j + B_{33} 1
\end{pmatrix}
= \matx B \begin{pmatrix}i\\j\\1\end{pmatrix}, \label{eqn:pinhole}
\end{equation}
where the $3\times3$ matrix $\matx B = \matx R^\top \matx K^{-1}$, and
$\matx R$ is often chosen to be the identity~\cite{Hartley00}.  The
rational function model handles lens distortion by permitting
$i$ and $j$ to appear in higher order polynomials, in particular
quadratic:
\begin{equation}
\vect d(i,j) = \begin{pmatrix}
A_{11} i^2 + A_{12} ij + A_{13} j^2 + A_{14} i + A_{15} j + A_{16}\\
A_{21} i^2 + A_{22} ij + A_{23} j^2 + A_{24} i + A_{25} j + A_{26}\\
A_{31} i^2 + A_{32} ij + A_{33} j^2 + A_{34} i + A_{35} j + A_{36}
           \end{pmatrix}.
           \label{eqn:model}
\end{equation}

\newcommand{\A}{\mbox{${\tt A}$}}
\newcommand{\Chi}{\raisebox{.4ex}{$\boldsymbol\chi$}}
This model may be written as a linear combination of the distortion
parameters, in a $3\times6$ matrix \A\ (analogous to $\matx B$
above), and a 6-vector $\Chi$ of monomials in $i$ and $j$.  Define
$\Chi$ as the ``lifting'' of image point $(i,j)$ to a six
dimensional space
\begin{equation}
\Chi(i,j) =
[i^2,\;\;ij,\;\;j^2,\;\;i,\;\;j,\;\;1]^{\top}\label{eqn:Chi}
\end{equation}
The imaging model~(\ref{eqn:model}) may then be written
\begin{equation}
\vect d(i,j) = \A \Chi(i,j) \label{eqn:project}
\end{equation}

\def\conic{{\boldsymbol\theta}}
A line in the scene forms a plane with the origin of camera
coordinates, and is imaged to the set of $\vect d$ in that plane.
This yields the line equation $\vect l\tr \vect d = 0$ which, in
terms of image points $(i,j)$ becomes
\begin{equation}
\vect l\tr \A \Chi = 0 \Longleftrightarrow \conic^{\top} \Chi = 0,
\end{equation}
where $\conic = (A_{xx}, A_{xy}, A_{yy}, A_x, A_y,
A_{_\mathit{0}})\tr$ are the parameters of a conic in image coordinates $(i,j)$
and $\Chi$ is given by (\ref{eqn:Chi}).  Here we observe the
important property that lines in the world go to conics under the
rational function model.  The task of calibration is then to find an
\A\ which will map these conics in the distorted image back to
straight lines.

By fitting a conic to the image of the line, we obtain
parameters~$\conic$, and thus the constraint
$\conic = \A\tr \vect l$
for unknown $\A$ and $\vect l$.  The equality is exact as any scale
factor is included in $\vect l$. Collecting $L$ such constraints, we
obtain
\begin{align}
\underbrace{\left[ \conic_1 \mid \ldots \mid \conic_L
\right]}_{6\times L} & =
\underbrace{\phantom{\left[\right]}\A\tr}_{6\times 3}
\underbrace{\left[ \vect l_1 \mid \ldots \mid \vect l_L \right]}_{3\times L}
\nonumber
\end{align}
which we write as
$\matx C = \A\tr \matx L\label{eqn:proj_line}$
so the matrix of conic parameters $\matx C$ is of rank no greater
than~$3$.  Therefore $\A$ can be computed up to a homography by
factorization: if $\matx U \matx S \matx V\tr = \matx C$ is the SVD
of $\matx C$, then $\A = \matx S_{(1:3,1:3)} {\matx U_{(:,1:3)}}\tr$
is one member of the equivalence class of solutions.

\begin{figure}[t]
\includegraphics[width=\linewidth]{images/fig1.png}
\caption{
Edges corresponding to straight lines in the real world are
detected (a) and the plumbline constraint is used to compute the
distortion parameters, giving the rectified image (b)}
\vspace{-2mm}
\end{figure}

The matrix $\matx C$ will not be rank 3 if the conics were obtained by
fitting to noisy image data.  The above factorization truncates $\matx C$
to rank 3 by minimizing the error in the conic parameter space, and as
discussed in the paper, this is sensitive to noise.  The strategy for
fitting \A\ from noisy image data is to run a non-linear optimization that
finds the \A\ which minimizes the error between the image data and straight
lines projected (as conics) into the distorted image.  The nonlinear
objective measures the Sampson distance~\cite{Hartley00} from the conics to
the detected edgels.  Let $\vect e_{\ell k} = (i_{\ell k},j_{\ell k})$
denote the $k^{th}$ edgel in linked segment $\ell$. The Sampson distance is
a first order approximation to the distance from a point to a conic. The
error function $\varepsilon(\matx A, \vect l_1, \ldots, \vect l_L)$ we
minimize is then given by (with $\conic^\ell := \matx A \vect l_\ell$)
\begin{equation}
\def\t#1{\conic^\ell_{#1}}
\varepsilon =
\sum_{\ell = 1}^{L} \sum_{k=1}^{n_\ell }
\frac
{\left[(\conic^\ell)\tr\Chi\left(i_{\ell k},j_{\ell k}\right)\right]^2}
{(2\t1 i_{\ell k} + \t2 j_{\ell k} + \t4)^2 + (2\t3 j_{\ell k} +
\t2 i_{\ell k} + \t5)^2}
\nonumber
\end{equation}
Implementation of this method by edge detection and edge linking is
described in the paper, as are the details of the nonlinear optimization.
Our conclusion is that the simplicity of the rational function model,
coupled with its ability to model a variety of lenses, makes it a useful
model to consider when dealing with lenses exhibiting moderate to severe
distortion.

>>>>

% Use class option [extendedabs] to prepare the 1-page extended abstract.
\documentclass[extendedabs]{bmvc2k}
\usepackage[colorlinks = true,
            linkcolor = blue,
            urlcolor  = blue,
            citecolor = blue,
            anchorcolor = blue]{hyperref}
\graphicspath{ {figures/} }

@@

\title{Enriching Object Detection with 2D-3D Registration and Continuous Viewpoint Estimation}
\addauthor{
Christopher Bongsoo Choy\textsuperscript{\dag}, Michael Stark\textsuperscript{\dag \ddag}, Sam Corbett-Davies\textsuperscript{\dag}, Silvio Savarese\textsuperscript{\dag}}{}{1}
\addinstitution{
\textsuperscript{\dag}Stanford University, \textsuperscript{\ddag}Max Planck Institute for Informatics
}

@@

\noindent
% In this paper, we propose a novel method for 2D-3D alignment of exemplar CAD
% models to real-world images that circumvents the need for calibration and
% greatly enhances the scalability of WHO. Our model generates a discriminative
% exemplar template on-the-fly and align model accurately using MCMC.

A large body of recent work on object detection has focused on providing more
than a bounding box of an object. Some exploited 3D CAD
model databases to provide viewpoint estimation. These approaches work
by aligning exact 3D models to images using templates generated from renderings
of the 3D models at a set of discrete viewpoints.

However, training the template for all viewpoint and CAD models is not feasible
since gathering all training data for different viewpoint and CAD models is
expensive.  Instead, it has been realized that template-based exemplar
detectors based on HOG features can be trained analytically, by
replacing the standard SVM with an LDA classifier $w_{x_t}$ for a template $x_t$
~\cite{Hariharan12}.

The result is a whitened feature representation, termed WHO (Whitened Histogram
of Orientations). This development makes it feasible to train hundreds of
thousands of mid-level patch detectors for recognition.

Though it achieved remarkable improvements, generating LDA templates (matrix
decomposition with Gaussian Elimination) and calibrating them are
computationally expensive and require gigabytes of memory and storage, while
the viewpoint discretization hampers pose estimation performance.

% In this paper, instead of caching and calibrating all templates on training
% time, we selectively generate templates on testing time as well to estimate
% viewpoint continuously using the technique we proposed termed Non-Zero Whitened
% Histograms of Orientations (NZ-WHO).

% The method uses Conjugate Gradient method to efficiently generate an LDA template on-the-fly (60 ms) without expensive calibration steps.

\begin{figure}[t]
  \centering
  % \includegraphics[width=0.9\linewidth]{schematics} 
  \includegraphics[width=0.8\linewidth]{front} % \\[-5pt]
  \caption{Using a database of 3D CAD models, we generate NZ-WHO
    templates which can be used to either detect objects directly or
    enrich the output of an existing detector with high-quality,
    continuous pose and 3D CAD model exemplar.}
    % initialize pose of an object in the bounding box from other
    % detectors. Once initialization is given, we use NZ-WHO to propose
    % and validate plausible pose on-the-fly and
    % further tune the translation, 3D viewpoint and focal length continuously.}
  % \vspace{-1em}
  \label{fig:front}
\end{figure}

\paragraph{Overview}

In this paper, we propose a novel method for 2D-3D alignment of exemplar CAD
models to real-world images that circumvents the need for calibration and
greatly enhances the scalability of WHO. As a result, we can render novel views
and train corresponding exemplar models {\em on-the-fly}, without the need for
offline processing. We call these Non-Zero Whitened Histograms of Orientations
(NZ-WHO) templates Fig.~\ref{fig:front}. As a by-product, we can formulate the
alignment problem as a parameter search in a continuous pose space, consisting
of yaw, pitch and roll, which we implement using MCMC sampling.

First, we present a novel method for training exemplar models from
rendered 3D CAD data on-the-fly, enabled by a novel variant of
WHO, termed Non-Zero Whitened Histograms of Orientations (NZ-WHO), and making
efficient use of the specific characteristics of rendered images. To our
knowledge, our method constitutes the first attempt to simultaneously render
and train exemplar detectors on-the-fly.

Second, we demonstrate that our method can enrich the output of an existing
object class detector, such as the DPM or the
R-CNN with additional 3D information. By applying our
method to candidate detections provided by the respective detector, we can
augment the original detections with an estimate of 3D continuous pose and a 3D
CAD model exemplar.

Finally, we give an in-depth experimental study that demonstrates
the effectiveness of our approach on a standard benchmark for
object detection and viewpoint estimation~\cite{Xiang14}.

\begin{figure}[t]
\small
\setlength\tabcolsep{1pt}
\centering
\begin{tabular}{ccccc}
  \includegraphics[width=0.2\textwidth]{car_cnn/2c.png} &
  \includegraphics[width=0.2\textwidth]{car_cnn/2e.png} \\
  \includegraphics[width=0.2\textwidth]{bicycle_cnn/4b.png} &
  \includegraphics[width=0.2\textwidth]{bicycle_cnn/4c.png} \\
\end{tabular}
\caption{Example enriched bounding boxes. Given R-CNN
	detection bounding boxes, our method predicts the best 2D-3D
	registration given the bounding box input. Blue boxes are R-CNN output and
	purple boxes are the tightest bounding box enclosing predicted CAD model.}
\label{fig:pascal12cnn}
\end{figure}

\paragraph{Non-Zero Whitened Histograms of Orientations}

NZ-WHO removes these artifacts so that the background has no effect on the
template response. Let a vectorized HOG feature of a rendering image $x = [
x_1^T ,  \cdots , x_n^T ]^T \in \mathcal{R}^{nd}$ where $x_i$ is the $i$th HOG cell feature, $n$ is the
number of HOG cells and $d$ is the dimension of the HOG feature. We create a new
vector $\bar{x}$ which contains only the non-zero HOG cell features of $x$.
Synthesizing the LDA template requires solving the system of linear equations
$\bar{w}=\bar{\Sigma}^{-1}(\bar{x} - \bar{\mu})$. In \cite{Hariharan12}, the
authors solve the system via the Cholesky decomposition, which requires
$O(n^3)$ time. Instead, we introduced the Conjugate Gradient method to generate
LDA templates. The method runs in $O(n^2\kappa)$ time, where $\kappa$ is the condition number of the matrix and achieved more than 100x speed-up.

% Also, the residual from the Conjugate Gradient method (the norm of $y-Ax$ for
% $Ax = y$) is smaller than 
% that of Cholesky decomposition.

\paragraph{Fine-Tuning using MCMC}
Due to the speed of our template synthesis procedure, we are able to perform
joint optimization of scale, translation, continuous rotation, and focal length
using the Metropolis-Hastings algorithm.

We model the probability of an object with the parameter $\theta$ in the test
image $\mathcal{I}$ as $P(\theta| \mathcal{I}) \sim e^{ \max_{s} w(\theta)
	\ast \mathcal{T}_s(\mathcal{I})}$. We approximate the MAP solution for $p$ by
drawing samples from the distribution $P(\theta | \mathcal{I})$, using the
Metropolis-Hastings algorithm.

\paragraph{Experiments}

First, we verify that our NZ-WHO method delivers performance that is
at least on par with the original WHO formulation~\cite{Hariharan12} in terms of
accuracy, while at the same time resulting in large computational
savings.

Second, we demonstrate that our method can be used for multi-view
object class detection in isolation. It can be applied in a sliding
window fashion and deliver 2D bounding box as well as viewpoint
information. Our method is competitive with the state-of-the-art in this
case.

Finally, we show that our method can be used to complement the
detections provided by an existing object class detector, such
as DPM or RCNN. In this case, we show a considerable performance improvement
compared to previous work in the task of joint object class detection and VP
estimation. Example outputs of our model are on Fig.~\ref{fig:pascal12cnn}.

### 2015-04-07 approach.tex
extended abstract and Michael's edit

We used an off-the-shelf rendering engine to generate a realistic rendering
and a depth map. The CAD models we used contain texture and material
information. These make the rendering more realistic and allow us to transfer
natural image statistics to rendered images. During rendering we extract a depth map which can be used for
reconstruction\ref{fig:rendering}.

@@

\subsection{Whitened Histograms of Orientations (WHO)}
\label{sec:who}
Our technique for rendering and generating exemplar template detectors
on-the-fly draws from recent work by Hariharan \etal~\cite{Hariharan12}. They introduced Whitened
Histograms of Orientations (WHO), which uses feature statistics from natural
images to create a large number of classifiers analytically using Linear
Discriminant Analysis (LDA) rather than training SVM
classifiers. The confidence score for data $x_i$ can be
defined as $ S(x_i) = w_{x_t}^T x_i$ where $w_{x_t} = \Sigma^{-1} (x_t - \mu)$
is an LDA classifier for a template $x_t$. Since collecting covariance matrices
for all possible template shape is intractable, \cite{Hariharan12} assumed
Wide-Sense Stationarity(WSS) of HOG features and generated a covariance
$\Sigma$ from autocovariance $\Gamma$ collected on a large collection of
natural images. For a 31 dim HOG feature $x_t$ at location $t$

\begin{align}
    E[x_t] & = \mu\\
    E[(x_{t} - \mu ) (x_{\tau} - \mu)] & = E[(x_0 - \mu)(x_{\tau - t} - \mu)].
\end{align}

In this paper, we followed their method to generate $\Sigma$.

@@

NZ-WHO removes these artifacts so that the background has no effect on the
template response. Let a vectorized HOG features of a rendering image $x = [
\begin{array}{ccccc}x_1^T & x_2^T & \cdots & x_n^T\end{array} ]^T \in
\mathcal{R}^{nd}$ where $x_i$ is the $i$th HOG cell feature, $n$ is the
number of HOG cells and $d$ is the dimension of the HOG feature. We create a new
vector $\bar{x}$ which contains only the non-zero HOG cell features of $x$. To be
specific, let $I_d\in \mathcal{R}^{d\times d}$ be the identity matrix, $\bar{n}$ be
the number of non-zero HOG cells, and a matrix $S\in \mathcal{R}^{\bar{n}d \times
    nd}$ be the masking matrix that selects non-zero HOG cells. For instance, a
template has  $n=3, \bar{n}=2$, and only the second HOG cell has $0$ norm, then

@@

\begin{equation}
    \bar{w}=\bar{\Sigma}^{-1}(\bar{x} - \bar{\mu}) \label{eq:nz-who},
\end{equation}

@@

\begin{figure}[t]
  \begin{center}
%   \fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
    % include whitening all centered cells
    % \includegraphics[width=0.32\linewidth]{whiten_all_crop}
    \setlength\tabcolsep{3pt}
    \begin{tabular}{ccc}
      HOG & WHO & NZ-WHO \\
%     \begin{turn}{90}$w_+$\end{turn} &
    \includegraphics[width=0.28\linewidth]{hog_crop} &
    \includegraphics[width=0.28\linewidth]{whiten_all_crop} &
    \includegraphics[width=0.28\linewidth]{whiten_non_zero_crop} \\
    % include whitening all centered cells
     % \includegraphics[width=0.282\linewidth]{whiten_all_neg_crop} 
%     \begin{turn}{90}$-w_-$\end{turn} &
     \includegraphics[width=0.28\linewidth]{hog_neg_crop} &
     \includegraphics[width=0.28\linewidth]{whiten_all_neg_crop}  &
     \includegraphics[width=0.28\linewidth]{whiten_non_zero_neg} \\
    % include whitening all centered cells
 %    \begin{turn}{90}ihog\cite{vondrick2013}\end{turn} &
    % \cite{vondrick2013}& 
     \includegraphics[width=0.28\linewidth]{ihog_hog200_crop.png} &
     \includegraphics[width=0.28\linewidth]{ihog_whiten_all200_crop.png} &
     \includegraphics[width=0.28\linewidth]{ihog_whiten_non_zero200_crop.png} \\
 \end{tabular}
  \end{center}
  \caption{Comparison of HOG, WHO and NZ-WHO. Visualization of positive weights (first row),  visualization of negative weights (second row), HOGgles \cite{vondrick2013} (third row). Note that for WHO, whitening all cells results in strong negative edges on the empty region}
  \label{fig:whocomparison}
\end{figure}

@@

As a result, a GPU implementation of conjugate gradient converges in 60
ms when using 250 HOG cells on Nvidia GTX660, two orders of magnitude faster
than the using Cholesky factorization with Gaussian Elimination.

We report the real time analysis of whitening using decomposition and
Conjugate Gradient methods in Fig.~\ref{fig:whotime}. (a) compares the
absolute runtime of the different methods while (b) gives the obtained
speedup. We see that % We compare the speed of each of template generation methods in
% Fig.~\cite{fig:covariancetime_crop} Using naive Cholesky decomposition,
whitening the HOG template takes several seconds for realistic template sizes of
several hundred cells, but only 60ms using the Conjugate Gradient
method. If we use NZ-WHO, we can gain extra speed
up since we only whiten non-zero cells.

@@

\begin{figure}[t]
  \begin{center}
  \begin{tabular}{cc}
     \includegraphics[width=0.5\linewidth]{whotime} & 
     \includegraphics[width=0.5\linewidth]{speedup}\\
     (a) & (b) \\
 \end{tabular}
  \end{center}
  \caption{(a) Runtime analysis of whitening. HOG means feature
    extraction time, WHO-Chol uses ~\cite{Hariharan12} and
    WHO-CG uses iterative Conjugate Gradient. NZ-WHO-CG (Ours) uses only
    non-zero cells and Conjugate Gradient. (b) final speedup
    of NZ-WHO vs. WHO.}
  \label{fig:whotime}
\end{figure}
%
\begin{figure}[t]
  \centering
  \includegraphics[width=0.5\linewidth]{residual}
  \caption{Residuals from different method}
  \label{fig:whoresidual}
\end{figure}

\subsection{High Resolution Templates and FFT based Convolution}
\label{sec:fft}

>>>>

We use an off-the-shelf rendering engine to generate a realistic rendering
and a depth map. The CAD models we used contain texture and material
information. These make the rendering more realistic and allow us to transfer
natural image statistics to rendered images Fig.~\ref{fig:rendering}.

@@

\subsection{Whitened Histograms of Orientations (WHO)}
\label{sec:who}
Our technique for rendering and generating exemplar template detectors
on-the-fly draws from recent work by Hariharan \etal~\cite{Hariharan12}. They introduced Whitened
Histograms of Orientations (WHO), which uses feature statistics from natural
images to create a large number of classifiers analytically using Linear
Discriminant Analysis (LDA) rather than training SVM
classifiers. The confidence score for data $x_i$ can be
defined as $ S(x_i) = w_{x_t}^T x_i$ where $w_{x_t} = \Sigma^{-1} (x_t - \mu)$
is an LDA classifier for a template $x_t$. Since collecting covariance matrices
for all possible template shape is intractable, \cite{Hariharan12} assumed
Wide-Sense Stationarity (WSS) of HOG features and generated a covariance
$\Sigma$ from autocovariance $\Gamma$ collected on a large collection of
natural images. In this paper, we further assumed symmetry of the
autocovariance. For a 31-dimension HOG feature $x_t$ at location $t$

\begin{align}
    E[x_t] & = E[x_\tau] = \mu\\
    \Gamma_{\|t - \tau\|} & = E[(x_{t} - \mu ) (x_{\tau} - \mu)] \\
              & = E[(x_0 - \mu)(x_{\tau - t} - \mu)].
\end{align}

for all $t$ and $\tau$. In practice, we gathered $\Gamma$ up to $|t - \tau |
\le 40$ both horizontally and diagonally.

@@

NZ-WHO removes these artifacts so that the background has no effect on the
template response. Let a vectorized HOG feature of a rendering image $x = [
\begin{array}{ccccc}x_1^T & x_2^T & \cdots & x_n^T\end{array} ]^T \in
\mathcal{R}^{nd}$ where $x_i$ is the $i$th HOG cell feature, $n$ is the
number of HOG cells and $d$ is the dimension of the HOG feature. We create a new
vector $\bar{x}$ which contains only the non-zero HOG cell features of $x$. To be
specific, let $I_d\in \mathcal{R}^{d\times d}$ be the identity matrix, $\bar{n}$ be
the number of non-zero HOG cells, and a matrix $S\in \mathcal{R}^{\bar{n}d \times
nd}$ be the masking matrix that selects non-zero HOG cells. For instance, a
template has  $n=3, \bar{n}=2$, and only the second HOG cell has $0$ norm, then

@@

\begin{equation}
    \bar{w}=\bar{\Sigma}^{-1}(\bar{x} - \bar{\mu}) \label{eq:nz-who}
\end{equation}

@@

\begin{figure}[t]
  \begin{center}
%   \fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
    % include whitening all centered cells
    % \includegraphics[width=0.32\linewidth]{whiten_all_crop}
    \setlength\tabcolsep{3pt}
    \begin{tabular}{ccc}
      HOG & WHO & NZ-WHO \\
%     \begin{turn}{90}$w_+$\end{turn} &
    \includegraphics[width=0.28\linewidth]{hog_crop} &
    \includegraphics[width=0.28\linewidth]{whiten_all_crop} &
    \includegraphics[width=0.28\linewidth]{whiten_non_zero_crop} \\
    % include whitening all centered cells
     % \includegraphics[width=0.282\linewidth]{whiten_all_neg_crop} 
%     \begin{turn}{90}$-w_-$\end{turn} &
     \includegraphics[width=0.28\linewidth]{hog_neg_crop} &
     \includegraphics[width=0.28\linewidth]{whiten_all_neg_crop}  &
     \includegraphics[width=0.28\linewidth]{whiten_non_zero_neg} \\
    % include whitening all centered cells
 %    \begin{turn}{90}ihog\cite{vondrick2013}\end{turn} &
    % \cite{vondrick2013}& 
     \includegraphics[width=0.28\linewidth]{ihog_hog200_crop.png} &
     \includegraphics[width=0.28\linewidth]{ihog_whiten_all200_crop.png} &
     \includegraphics[width=0.28\linewidth]{ihog_whiten_non_zero200_crop.png} \\
 \end{tabular}
  \end{center}
  \caption{Comparison of HOG, WHO and NZ-WHO. Visualization of positive weights
      (first row),  visualization of negative weights (second row), HOGgles
      \cite{vondrick2013} (third row). Note that for WHO, whitening all cells
      results in strong negative edges on the empty region.}
  \label{fig:whocomparison}
\end{figure}

@@

As a result, a GPU implementation of conjugate gradient converges in 60
ms when using 250 HOG cells on Nvidia GTX660, two orders of magnitude faster
than using Cholesky factorization with Gaussian Elimination.

We report the real time analysis of whitening using decomposition and
Conjugate Gradient methods in Fig.~\ref{fig:whotime}. (a) compares the
absolute runtime of the different methods while (b) gives the obtained
speedup. We see that % We compare the speed of each of template generation methods in
% Fig.~\cite{fig:covariancetime_crop} Using naive Cholesky decomposition,
whitening the HOG template takes several seconds for realistic template sizes of
several hundred cells, but only 60 ms using the Conjugate Gradient
method. If we use NZ-WHO, we can gain extra speed
up since we only whiten non-zero cells.

@@

\begin{figure}[t]
  \begin{center}
  \begin{tabular}{cc}
     \includegraphics[width=0.5\linewidth]{whotime} & 
     \includegraphics[width=0.5\linewidth]{speedup}\\
     (a) & (b) \\
 \end{tabular}
  \end{center}
  \caption{(a) Runtime analysis of whitening. HOG means feature
    extraction time, WHO-Chol uses our implementation of ~\cite{Hariharan12} and
    WHO-CG uses iterative Conjugate Gradient. NZ-WHO-CG (Ours) uses only
    non-zero cells and Conjugate Gradient. (b) final speedup
    of NZ-WHO vs. WHO.}
  \label{fig:whotime}
\end{figure}
%
\begin{figure}[t]
  \centering
  \includegraphics[width=0.5\linewidth]{residual}
  \caption{Residuals of different method}
  \label{fig:whoresidual}
\end{figure}

\subsection{High Resolution Templates and FFT-based Convolution}
\label{sec:fft}
